---
output: github_document
markdown_extensions:
    - admonition
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# [U]{.underline}nited [S]{.underline}tates [En]{.underline}ergy [Sys]{.underline}tem Optimization Model

::: warn
> :warning: **This package is under development.**  
> The expected release date of the first minimal *beta*-version is 2024-09-30.
> Docs in progress: <https://usensys.github.io/usensys/>
:::

**USENSYS** is a linear programming optimization model of the US energy system. This class of models is known as **Macro Energy Models** or **Reference Energy System Models** or **Capacity Expansion Models** (for electric power sector) optimize capacity expansion of alternative technologies under alternative resource and policy constraints and assumptions.   
**usensys** r-package is an R-interface to the model with a set of functions to run the model, analyze the results, and visualize the data. The model is designed to be used for policy analysis, scenario analysis, and long-term planning of the energy system. The model is based on the open-source R-package **energyRt** and is designed to be used in combination with other R-packages for data analysis, visualization, and optimization. 


<!-- badges: start -->

<!-- badges: end -->

## Versions
Several versions of the model are available with different status of development, all will be included in the package. The versions differ in the geographical scope of the model and regional aggregation, the data sources, and/or replication of existing models by EPA, NREL, and others. The current coverage of economic sectors is currently focused on the electric power sector, but the model is designed to be extended to other sectors in the future. The following versions are currently in preparation for the release:

- **26-zones**: The model is based on 26 zones (NERC subscriptions) of the United States. This version has been a part of the Multimodel Intercomparison Project (MIP) and is based on the PowerGenome data.
- **50-states**: The model is based on the 50 states of the United States, including Alaska and Hawaii. The datasets are based on the EIA data and the NREL ATB. This regional division might be helpful for for integration with other sectors and for state-level policy analysis.
- **IPM**: The model is based on the IPM model by EPA. The datasets are based on the IPM data and the NREL ATB. 
- **IMMX**: The model is an extension of the IPM model by EPA with added Mexican states. 
- **ReEDS**: The model is a replica of ReEDS model by NREL with imported data from the ReEDS model repository.


## Prerequisites
One of the following optimization software is required to run the model:
- Julia with JuMP
- Python with Pyomo
- GAMS
- GLPK/MathProg (for small problems only)

Large problems require a commercial solver, such as CPLEX or Gurobi. Smoler models can be solved with open-source solvers, such as HiGSS, CBC, or GLPK.

## Installation

You can install the development version of usensys from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pkg_install("energyRt/energyRt")
pak::pkg_install("usensys/usensys")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(usensys)
## basic example code
```

:construction: **This document is under construction.** :construction:
