---
title: "Data for *usensys* models"
author: '@olugovoy'
date: "June 13, 2019"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    highlight: tango
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  results = "hide"
)
# knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)

library(readxl)
library(stringr)
library(rvest)
library(sp)
library(tidyverse)


```

# Info

The purpose of the code below is to acquire and preprocess US electric power sector data from EIA, EPA, and other sources. The final processed data, which is required to run USENSYS model, is stored in the '/data' directory in '.RData' format in the model working directory. It is not required to run scripts in this file to be able to run all the versions of USENSYS, available on GitHub. The code below is provided for reference, and in order to be able to change the aggregation procedure.

# Obtaining EIA data 
## By states

Downloading data from <https://www.eia.gov/electricity/data/state/>.  

Version:  
Final annual data for 2018
  Release Date: October 22, 2019
  Next Release Date: November 2020
  
```{r EIA_data_annual_&_monthly}
eiadir <- file.path(getwd(), "data/EIA")
if(!dir.exists(eiadir)) dir.create(eiadir, recursive = T)

{www <- list()
# Annual data ####
# 1990 - 2018	Net Generation by State by Type of Producer by Energy Source (EIA-906, EIA-920, and EIA-923)
www$annual_generation_state.xls <- "https://www.eia.gov/electricity/data/state/annual_generation_state.xls"
# 1990 - 2018	Fossil Fuel Consumption for Electricity Generation by Year, Industry Type and State (EIA-906, EIA-920, and EIA-923)2
www$annual_consumption_state.xls <- "https://www.eia.gov/electricity/data/state/annual_consumption_state.xls"
# 1990 - 2018	Existing Nameplate and Net Summer Capacity by Energy Source, Producer Type and State (EIA-860)1, 3
www$existcapacity_annual.xlsx <- "https://www.eia.gov/electricity/data/state/existcapacity_annual.xlsx"
# 2018 - 2023	Proposed Nameplate and Net Summer Capacity by Year, Energy Source, and State (EIA-860)1
www$plancapacity_annual.xlsx <- "https://www.eia.gov/electricity/data/state/plancapacity_annual.xlsx"
# 1990 - 2018	U.S. Electric Power Industry Estimated Emissions by State (EIA-767, EIA-906, EIA-920, and EIA-923)4
www$emission_annual.xls <- "https://www.eia.gov/electricity/data/state/emission_annual.xls"
# 1990 - 2018	Average Price by State by Provider (EIA-861)5
www$avgprice_annual.xlsx <- "https://www.eia.gov/electricity/data/state/avgprice_annual.xlsx"
# 1990 - 2018	Number of Retail Customers by State by Sector (EIA-861)5
www$customers_annual.xlsx <- "https://www.eia.gov/electricity/data/state/customers_annual.xlsx"
# 1990 - 2018	Retail Sales of Electricity by State by Sector by Provider (EIA-861)5
www$sales_annual.xlsx <- "https://www.eia.gov/electricity/data/state/sales_annual.xlsx"
# 1990 - 2018	Revenue from Retail Sales of Electricity by State by Sector by Provider (EIA-861)5
www$revenue_annual.xlsx <- "https://www.eia.gov/electricity/data/state/revenue_annual.xlsx"
# 2001 - 2003	Financial Data on Publicly Owned Electric Utilities with Generation Facilities by State (EIA-412)6	
www$financewgen_annual.xls <- "https://www.eia.gov/electricity/data/state/financewgen_annual.xls"
# 2001 - 2003	Financial Data on Publicly Owned Electric Utilities without Generation Facilities by State (EIA-412)6	
www$financewogen_annual.xls <- "https://www.eia.gov/electricity/data/state/financewogen_annual.xls"

# Monthly data from Electric Power Monthly ####
# Electric Power Monthly with data for November 2019
#   Release date: January 30, 2020
#   Next release date: End of February 2020
  
# 2001 - Present	Net Generation by State by Type of Producer by Energy Source1	
www$generation_monthly.xlsx <- "https://www.eia.gov/electricity/data/state/generation_monthly.xlsx"
# 2001 - Present	Fossil Fuel Consumption for Electricity Generation by Year, Industry Type and State2	
www$consumption_monthly.xlsx <- "https://www.eia.gov/electricity/data/state/consumption_monthly.xlsx"}

download <- function(url, path = eiadir, overwrite = FALSE) {
  fl <- file.path(path, basename(url))
  if(file.exists(fl) & !overwrite) {
    # message("File '", basename(url), "' already exists.")
    return("exists")
  } else {
    download.file(url = url, destfile = fl, mode = "wb") 
    if(file.exists(fl)) {
      return("downloaded")
    } else {
      return("download failed")
    }
  }
}

sapply(www, download, overwrite = F)

# Annual capacity
elc_cap <- read_excel(file.path(eiadir, basename(www$existcapacity_annual.xlsx)),
                      sheet = "Existing Capacity",
                      range = "A2:H44621")
elc_cap
       
# Fuels consumption for electricity generation by months
fue4elc <- read_excel(file.path(eiadir, basename(www$consumption_monthly.xlsx)),
                      sheet = "2018_Preliminary",
                      range = "A5:F9361")
fue4elc

# Electricity generation by months and energy type
elc_gen <- read_excel(file.path(eiadir, basename(www$generation_monthly.xlsx)),
                      sheet = "2018_Preliminary",
                      range = "A5:F25156")
elc_gen

elc_emis <- read_excel(file.path(eiadir, basename(www$emission_annual.xls)),
                      sheet = "State Emissions",
                      range = "A1:G41765")
elc_emis

save(elc_cap, fue4elc, elc_gen, elc_emis, file = file.path(eiadir, "eia_raw.RData"))

```


## Power plants & generator level data  
### Form EIA-860 detailed data with previous form data (EIA-860A/860B)  
<https://www.eia.gov/electricity/data/eia860/> 
  Release date: September 3, 2019, Final 2018 data
  Next release date: June 2020, Early release 2019 data  
The survey Form EIA-860 collects generator-level specific information about existing and planned generators and associated environmental equipment at electric power plants with 1 megawatt or greater of combined nameplate capacity.  

### Form EIA-923 detailed data with previous form data (EIA-906/920)  
<https://www.eia.gov/electricity/data/eia923/>  
Monthly (M) release date: January 30, 2020 for November 2019 data
  Next monthly release: End of February 2020 (December 2019 data)
  Annual release date: September 20, 2019, Final 2018 data
  2018 Re-released: January 8, 2020  Corrections/Revisions  

The survey Form EIA-923 collects detailed electric power data -- monthly and annually -- on electricity generation, fuel consumption, fossil fuel stocks, and receipts at the power plant and prime mover level. Specific survey information provided:
* Schedule 2 - fuel receipts and costs  
* Schedules 3A & 5A - generator data including generation, fuel consumption and stocks  
* Schedule 4 - fossil fuel stocks  
* Schedules 6 & 7 - non-utility source and disposition of electricity  
* Schedules 8A-F - environmental data  
Monthly data (M) -approximately 2,026 plants from the monthly survey  
Annual final data - approximately 2,026 monthly plants + 7,278 plants from the annual survey   

Summary data  
<https://www.eia.gov/electricity/data.php#gencapacity>  

### Scraping & downloading all zip and xls files of the forms
(Another option is EIA's API <https://www.eia.gov/opendata/register.php>, though it is not clear yet if the API has all the the data in xls, need to check.)  
```{r EIA_860_923_forms}
scrap <- function(weblink, filepatern = "([a-zA-Z0-9_]+.zip)|([a-zA-Z0-9_]+.xls)$", aslist = T) {
  # library(rvest)
  links <- read_html(weblink) %>%
    html_nodes("a") %>%
    html_attr('href') 
  links <- links[grepl(filepatern, links)]
  links <- paste0(weblink, links)
  if (aslist) {
    lnames <- str_extract(links, filepatern)
    stopifnot(length(links) == length(lnames))
    links <- as.list(links)
    names(links) <- lnames
  }
  links
}

www <- scrap("https://www.eia.gov/electricity/data/eia860/")
www <- c(www, scrap("https://www.eia.gov/electricity/data/eia923/"))
sapply(www, download, overwrite = F)

fls <- names(www)

# Select data for 2018
fls <- fls[grepl("2018", fls)]
zz <- grepl("zip$", fls) # select zip-files for unzipping
# Unzip files
for (f in fls[zz]) {
  dirname <- gsub("\\.zip$", "", f)
  dirname <- file.path(eiadir, dirname)
  if (!dir.exists(dirname)) dir.create(dirname)
  unzip(file.path(eiadir, f), exdir = dirname, overwrite = T)
}

# dirname <- gsub("\\.zip$", "", fls[1])
dbl2int <- function(x) {
  # coersing unneccesary doubles into integers
  stopifnot(any(class(x) == "data.frame"))
  ii <- sapply(x, is.double)
  # browser()
  for (i in (1:dim(x)[2])[ii]) {
    if (sum(x[[i]] - floor(x[[i]]), na.rm = T) == 0) x[[i]] <- as.integer(x[[i]])
  }
  x
}

YN2binary <- function(x) {
  # coersing columnds with Y or N data into logical
  stopifnot(any(class(x) == "data.frame"))
  ii <- sapply(x, is.character)
  # browser()
  for (i in (1:dim(x)[2])[ii]) {
    # Check values
    uval <- unique(x[[i]])
    if (length(uval) <= 4) {
      if (all(toupper(uval) %in% c("N","Y","X",NA))) {
        # browser()
        newx <- rep(NA, length(x[[i]]))
        newx[x[[i]] == "Y"] <- TRUE
        newx[x[[i]] == "N"] <- FALSE
        newx[x[[i]] == "X"] <- NA
        x[i] <- as.logical(newx)
      }
    }
  }
  x
}

```

Reading xls(x) files  
Note: reasing from Excel files should be supervised and adjusted in case of any updates of the files on EIA website. Also may not work as expected on some OS/platforms.
```{r}
cleanread_xls <- function(path, sheet, range, dbl2int = T, YNasBinary = T) {
  # It was obcerved, that original `read_excel` function may confuse data formats.
  # The 'cleanread_xls' function tries to fix this importing problem by
  # checking if any columns have been (mistakenly) imported as binary variables,
  # redeading the data with forced format.
  # Additional options/parameters:
  # dbl2int - convert numeric columns with no decimal part into integer
  # YNasBinary - convert columns with only "Y", "N", and <NA> values into logical
  xls <- read_excel(path, sheet, range)
  col_type <- sapply(xls, class)
  # browser()
  if (any(col_type == "logical")) { # reimport
    message("Rereading the xls-file, forcing out logical columns.")
    col_type[col_type == "character"] <- "text"
    col_type[col_type == "logical"] <- "text"
    xls <- read_excel(path, sheet, range, col_types = col_type)
  }
  if (dbl2int) xls <- dbl2int(xls)
  if (YNasBinary) xls <- YN2binary(xls)
  xls
}

elc_pp <- cleanread_xls(file.path(eiadir, "eia8602018/2___Plant_Y2018.xlsx"),
                      sheet = "Plant",
                      range = "A2:AP10982")
col_type <- sapply(elc_pp, class)
unique(col_type)
names(elc_pp)
unique(elc_pp$`Liquefied Natural Gas Storage`)

```

```{r, results="markdown", fig.cap="Mapping all power plants."}
load("data/maps/usa49reg.RData")
plot(usa49reg)
points(elc_pp$Longitude, elc_pp$Latitude, pch = 16, col = "blue", cex = .3)
save(elc_pp, file = file.path(eiadir, "elc_pp.RData"))

```


```{r}
# continue..
elc_gn <- cleanread_xls(file.path(eiadir, "eia8602018/3_1_Generator_Y2018.xlsx"),
                      sheet = "Operable",
                      range = "A2:BU22120")
col_type <- sapply(elc_gn, class)
unique(col_type)
elc_gn
names(elc_gn)

elc_genfue <- cleanread_xls(file.path(eiadir, "f923_2018/EIA923_Schedules_2_3_4_5_M_12_2018_Final_Revision.xlsx"),
                      sheet = "Page 1 Generation and Fuel Data",
                      range = "A6:CS13968")
elc_genfue
names(elc_genfue)
col_type <- sapply(elc_genfue, class)
unique(col_type)

# Split annual and monthly data
month_cols <- grepl(paste(month.name, collapse = "|"), names(elc_genfue))
summary(month_cols)

elc_genfue_y <- elc_genfue[, !month_cols]
dim(elc_genfue_y)

jj <- grepl("(Plant.id)|(YEAR)", names(elc_genfue), ignore.case = T)
elc_genfue_m <- bind_cols(elc_genfue[, which(jj)], elc_genfue[, month_cols])
dim(elc_genfue_m)

# reshape monthly data
aa <- gather(elc_genfue_m, key = "key", value = "value", -`Plant Id`, -YEAR)
aa$variable <- str_extract(aa$key, "^[a-zA-Z0-9_]+")
unique(aa$variable)
aa$month <- str_extract(aa$key, "[a-zA-Z0-9_]+$")
unique(aa$month)

elc_genfue_m <- select(aa, `Plant Id`, YEAR, month, variable, value)

elc_genfue <- left_join(elc_genfue_m, select(elc_genfue_y, -starts_with("Reserved")))

clean_names <- function(x) {
  # Replace special caracters in names of the data columns
  nx <- names(x) %>%
    str_replace_all("\r\n", "_") %>%
    str_replace_all("\r", "_") %>%
    str_replace_all("\n", "_") %>%
    str_replace_all(" ", "_") %>%
    str_replace_all("__", "_") %>%
    str_replace_all("\\$", "USD") %>%
    str_replace_all("/", "_") %>%
    str_replace_all("\\?", "") %>%
    str_replace_all("\\(", "") %>%
    str_replace_all("\\)", "")
  names(x) <- nx
  x
}

elc_genfue <- clean_names(elc_genfue)
names(elc_genfue)

```


### Processing plant-level data
```{r, eval=FALSE}
# Fuels
elc_genfue_y <- clean_names(elc_genfue_y)
names(elc_genfue_y)

unique(elc_genfue_y$Reported_Fuel_Type_Code)
unique(elc_genfue_y$AER_Fuel_Type_Code)

unique(elc_gn$Technology)
elc_gn <- clean_names(elc_gn)
names(elc_gn)

aa <- elc_gn %>% group_by(Technology, State) %>%
        summarise(cap = sum(`Nameplate_Capacity_MW`))

summary(aa)
sum(aa$cap)/1e3

(unique(aa$Technology))

```

# EPA data

## EPA's Power Sector Modeling Platform v6 - November 2018 Reference Case

"March 21, 2019 - EPA is making the latest power sector modeling platform available, including the associated input data and modeling assumptions, outputs, and documentation." <https://www.epa.gov/airmarkets/documentation-epas-power-sector-modeling-platform-v6-november-2018-reference-case>
Accessed: June 13, 2019

### Get IPM data files
```{r}
epadir <- file.path(getwd(), "data/EPA")
ipmdir <- file.path(epadir, "IPM")
# needsdir <- file.path(epadir, "IPM")

if(!dir.exists(ipmdir)) dir.create(ipmdir, recursive = T)

ipmweb <- "https://www.epa.gov/airmarkets/documentation-epas-power-sector-modeling-platform-v6-november-2018-reference-case"

scrap_ipm <- function(weblink, filepatern = "([a-zA-Z0-9_]+.zip)|([a-zA-Z0-9_]+.xls(|x))$", aslist = T) {
  # library(rvest)
  links <- read_html(weblink) %>%
    html_nodes("body") %>% html_nodes("section") %>% 
    html_nodes("div") %>% html_nodes("div") %>% 
    html_nodes("div") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes('span') %>% 
    html_nodes('a') %>% html_attr('href') 
  links <- links[grepl(filepatern, links)]
  links <- paste0("https://www.epa.gov", links)
  if (aslist) {
    lnames <- str_extract(links, filepatern)
    stopifnot(length(links) == length(lnames))
    links <- as.list(links)
    names(links) <- lnames
  }
  links
}
www <- scrap_ipm(ipmweb)
sapply(www, download, overwrite = F, path = ipmdir)

# IPM shape files
download(url = "https://www.epa.gov/sites/production/files/2019-08/ipm_v6_regions.zip",
         overwrite = F, path = ipmdir)
unzip(file.path(ipmdir, "ipm_v6_regions.zip"), 
      exdir = file.path(ipmdir, "ipm_v6_regions"), overwrite = T)

```


### Coal
```{r read_coal_data}
# Documentation
download(url = "https://www.epa.gov/sites/production/files/2019-03/documents/chapter_7.pdf", 
         path = ipmdir, overwrite = F)


# Read coal supply data
coa_sup <- cleanread_xls(file.path(ipmdir, "table_7-26_coal_supply_curves_in_epa_platform_v6.xlsx"),
                      sheet = "Table 7-26 Coal",
                      range = "B4:I9002")
col_type <- sapply(coa_sup, class)
unique(col_type)

coa_sup <- clean_names(coa_sup)
names(coa_sup)
coa_sup

unique(coa_sup$Coal_Grade)
unique(coa_sup$Coal_Supply_Region)
unique(coa_sup$Year)

# Coal type
coa_sup$Coal_Type <- substr(coa_sup$Coal_Grade, 1, 1)
unique(coa_sup$Coal_Type)

# Aggregate by coal types
coa_sup_agg <- coa_sup %>%
  group_by(Year, Coal_Type, Coal_Supply_Region) %>%
  summarise(
    Heat_Content_MMBtu_Ton = mean(Heat_Content_MMBtu_Ton, na.rm = T),
    Cost_of_Production_2016USD_Ton = 
      weighted.mean(Cost_of_Production_2016USD_Ton, 
                    Coal_Production_Million_Tons_Year, na.rm = T),
    Coal_Production_Million_Tons_Year = sum(Coal_Production_Million_Tons_Year, na.rm = T),
    Coal_Reserves_Million_Tons = sum(Coal_Reserves_Million_Tons, na.rm = T)
  )

coa_type_names <- c(B = "Bituminous", S = "Subbituminous", L = "Lignite")
coa_sup_agg$Coal_Type_Long <- plyr::revalue(coa_sup_agg$Coal_Type, coa_type_names)
# write_csv(coa_sup_agg, path = "tmp/coa_sup_agg.csv")
write_tsv(coa_sup_agg, path = "tmp/coa_sup_agg.tab")

```

### Gas
```{r read_gas_Data}
download(url = "https://www.epa.gov/sites/production/files/2019-03/documents/chapter_8.pdf", 
         path = ipmdir, overwrite = F)

# Read gas supply curves data
gas_sup_curves <- cleanread_xls(
  file.path(ipmdir, "table_8-5_natural_gas_supply_curves_for_epa_platform_v6.xlsx"),
  sheet = 1,
  range = "B4:D484")

gas_sup_curves <- clean_names(gas_sup_curves)

# Aggregated supply with average prices
gas_sup_avr <- gas_sup_curves %>%
  group_by(Year) %>%
  summarise(Price_2016USD = weighted.mean(Price_2016USD, Natural_Gas_Supply_to_Power_Sector_Quads),
            Natural_Gas_Supply_to_Power_Sector_Quads = sum(Natural_Gas_Supply_to_Power_Sector_Quads))

# Read basis data
gas_sup_basis <- cleanread_xls(
  file.path(ipmdir, "table_8-4_natural_gas_basis_for_epa_platform_v6.xlsx"),
  sheet = 1,
  range = "B4:F596")
gas_sup_basis <- clean_names(gas_sup_basis)

all(unique(gas_sup_basis$Year) == unique(gas_sup_avr$Year)) # Check

# Merge with prices
gas_sup_byreg <- full_join(gas_sup_basis, gas_sup_avr)
unique(gas_sup_byreg$IPM_Region)

write_tsv(gas_sup_avr, path = "tmp/gas_sup_avr.tab")
write_tsv(gas_sup_byreg, path = "tmp/gas_sup_byreg.tab")


```


### NEEDS v6 Database
National Electric Energy Data System (NEEDS) v6
<https://www.epa.gov/airmarkets/national-electric-energy-data-system-needs-v6>
Accessed: Feb 13, 2020

The National Electric Energy Data System or “NEEDS” database contains the generation unit records used to construct the "model" plants that represent existing and planned/committed units in EPA modeling applications of IPM. NEEDS includes basic geographic, operating, air emissions, and other data on these generating units. NEEDS v6 is a complete bottom-up update for EPA's new power sector modeling platform v6.  For a description of the sources used in preparing NEEDS v6, see Documentation, Chapter 4: Generating Resources.


```{r}
needsweb <- "https://www.epa.gov/airmarkets/national-electric-energy-data-system-needs-v6"

# Most Recent Version of NEEDS:
#  NEEDS v6 rev: 12-18-2019(5 MB)
# Comment and Change Log rev: 12-18-2019(128 K)
download(url = "https://www.epa.gov/sites/production/files/2019-12/needs_v6_12-18-2019.xlsx", 
         path = ipmdir, overwrite = F)


# NEEDS v6 Database(4 MB)  (rev: 5-31-2019)
# download(url = "https://www.epa.gov/sites/production/files/2019-06/needs_v6_may_2019_reference_case.xlsx", 
#          path = ipmdir, overwrite = F)


# Comment and Change Log(88 K) (rev: 5-31-2019)

# NEEDS v6 Database(3 MB)  (rev: 11-30-2018)
# download(url = "https://www.epa.gov/sites/production/files/2019-03/needs_v6_november_2018_reference_case_0.xlsx", 
#          path = ipmdir, overwrite = F)

# Incremental Documentation for NOx Rates in NEEDS v6(3 pp, 44 K) (9-14-2018)

# NEEDS v6 Database(3 MB) (rev: 5-30-2018)


scrap_needs <- function(weblink, filepatern = "([a-zA-Z0-9_]+.zip)|([a-zA-Z0-9_]+.xls(|x))$", aslist = T) {
  # library(rvest)
  links <- read_html(weblink) %>%
    html_nodes("body") %>% html_nodes("section") %>% 
    html_nodes("div") %>% html_nodes("div") %>% 
    html_nodes("div") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes('span') %>% 
    html_nodes('a') %>% html_attr('href') 
  links <- links[grepl(filepatern, links)]
  links <- paste0("https://www.epa.gov", links)
  if (aslist) {
    lnames <- str_extract(links, filepatern)
    stopifnot(length(links) == length(lnames))
    links <- as.list(links)
    names(links) <- lnames
  }
  links
}
www <- scrap_ipm(needsweb)
sapply(www, download, overwrite = F, path = ipmdir)


```

### eGRID
...

