---
title: "Data for *usensys* models"
author: '@olugovoy'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    highlight: tango
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  results = "hide"
)
# knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)

library(readxl)
library(stringr)
library(rvest)
library(sp)
library(tidyverse)

# Function to download
download <- function(url, path = eiadir, overwrite = FALSE) {
  fl <- file.path(path, basename(url))
  if(file.exists(fl) & !overwrite) {
    # message("File '", basename(url), "' already exists.")
    return("exists")
  } else {
    download.file(url = url, destfile = fl, mode = "wb") 
    if(file.exists(fl)) {
      return("downloaded")
    } else {
      return("download failed")
    }
  }
}

cleanread_xls <- function(path, sheet, range = NULL, skip = NULL,
                          dbl2int = T, YNasBinary = T) {
  # It was observed, that original `read_excel` function may confuse data formats.
  # The 'cleanread_xls' function tries to fix this importing problem by
  # checking if any columns have been (mistakenly) imported as binary variables,
  # and rereading the data with forcing new guessed format.
  # Additional options/parameters:
  # dbl2int - converts numeric columns with no decimal part into integer
  # YNasBinary - converts columns with only "Y", "N", and <NA> values into logical
  cat("Reading the xls-file. ")
  xls <- suppressWarnings(read_excel(path, sheet, range, skip = skip))
  col_type <- sapply(xls, class)
  # browser()
  cat("Checking types of columns.")
  if (any(col_type == "logical")) { # reimport
    cat("\n")
    message("Rereading the file, correcting types of columns.")
    col_type[col_type == "character"] <- "text"
    col_type[col_type == "logical"] <- "text"
    xls <- read_excel(path, sheet, range, skip = skip, col_types = col_type)
  } else {
    cat(" All OK. ")
  }
  cat("done.")
  if (dbl2int) xls <- dbl2int(xls)
  if (YNasBinary) xls <- YN2binary(xls)
  xls
}

dbl2int <- function(x) {
  # coersing unneccesary doubles into integers
  stopifnot(any(class(x) == "data.frame"))
  ii <- sapply(x, is.double)
  # browser()
  for (i in (1:dim(x)[2])[ii]) {
    if (sum(x[[i]] - floor(x[[i]]), na.rm = T) == 0) x[[i]] <- as.integer(x[[i]])
  }
  x
}

YN2binary <- function(x) {
  # coersing columnds with "Y" and "N" values into logical
  stopifnot(any(class(x) == "data.frame"))
  ii <- sapply(x, is.character)
  # browser()
  for (i in (1:dim(x)[2])[ii]) {
    # Check values
    uval <- unique(x[[i]])
    if (length(uval) <= 4) {
      if (all(toupper(uval) %in% c("N","Y","X",NA))) {
        # browser()
        newx <- rep(NA, length(x[[i]]))
        newx[x[[i]] == "Y"] <- TRUE
        newx[x[[i]] == "N"] <- FALSE
        newx[x[[i]] == "X"] <- NA
        x[i] <- as.logical(newx)
      }
    }
  }
  x
}

clean_names <- function(x) {
  # Replaces special caracters in names of x
  nx <- names(x) %>%
    str_replace_all("\r\n", "_") %>%
    str_replace_all("\r", "_") %>%
    str_replace_all("\n", "_") %>%
    str_replace_all(" ", "_") %>%
    str_replace_all("__", "_") %>%
    str_replace_all("\\$", "USD") %>%
    str_replace_all("/", "_") %>%
    str_replace_all("\\?", "") %>%
    str_replace_all("\\(", "") %>%
    str_replace_all("\\)", "")
  names(x) <- nx
  x
}


```

# Info

The purpose of the code below is to acquire and preprocess US electric power sector data from EIA, EPA, and other sources. The final processed data, which is required to run USENSYS model, is stored in the '/data' directory in '.RData' format in the model working directory. It is not required to run scripts in this file to be able to run all the versions of USENSYS, available on GitHub. The code below is provided for the reference.

## Changelog
**April, 2020**   
- pre-processing data for the USENSYS-IPM  
- updated download/read functions  

...

**June 13, 2019**  
- initical commit, functions to webscrapp, download, and preprocess the data.  

# Obtaining EIA data 
## By states

Downloading data from <https://www.eia.gov/electricity/data/state/>.  

Version:  
Final annual data for 2018
  Release Date: October 22, 2019
  Next Release Date: November 2020
  
```{r EIA_data_annual_&_monthly}
eiadir <- file.path(getwd(), "data/EIA")
if(!dir.exists(eiadir)) dir.create(eiadir, recursive = T)

{www <- list()
# Annual data ####
# 1990 - 2018	Net Generation by State by Type of Producer by Energy Source (EIA-906, EIA-920, and EIA-923)
www$annual_generation_state.xls <- "https://www.eia.gov/electricity/data/state/annual_generation_state.xls"
# 1990 - 2018	Fossil Fuel Consumption for Electricity Generation by Year, Industry Type and State (EIA-906, EIA-920, and EIA-923)2
www$annual_consumption_state.xls <- "https://www.eia.gov/electricity/data/state/annual_consumption_state.xls"
# 1990 - 2018	Existing Nameplate and Net Summer Capacity by Energy Source, Producer Type and State (EIA-860)1, 3
www$existcapacity_annual.xlsx <- "https://www.eia.gov/electricity/data/state/existcapacity_annual.xlsx"
# 2018 - 2023	Proposed Nameplate and Net Summer Capacity by Year, Energy Source, and State (EIA-860)1
www$plancapacity_annual.xlsx <- "https://www.eia.gov/electricity/data/state/plancapacity_annual.xlsx"
# 1990 - 2018	U.S. Electric Power Industry Estimated Emissions by State (EIA-767, EIA-906, EIA-920, and EIA-923)4
www$emission_annual.xls <- "https://www.eia.gov/electricity/data/state/emission_annual.xls"
# 1990 - 2018	Average Price by State by Provider (EIA-861)5
www$avgprice_annual.xlsx <- "https://www.eia.gov/electricity/data/state/avgprice_annual.xlsx"
# 1990 - 2018	Number of Retail Customers by State by Sector (EIA-861)5
www$customers_annual.xlsx <- "https://www.eia.gov/electricity/data/state/customers_annual.xlsx"
# 1990 - 2018	Retail Sales of Electricity by State by Sector by Provider (EIA-861)5
www$sales_annual.xlsx <- "https://www.eia.gov/electricity/data/state/sales_annual.xlsx"
# 1990 - 2018	Revenue from Retail Sales of Electricity by State by Sector by Provider (EIA-861)5
www$revenue_annual.xlsx <- "https://www.eia.gov/electricity/data/state/revenue_annual.xlsx"
# 2001 - 2003	Financial Data on Publicly Owned Electric Utilities with Generation Facilities by State (EIA-412)6	
www$financewgen_annual.xls <- "https://www.eia.gov/electricity/data/state/financewgen_annual.xls"
# 2001 - 2003	Financial Data on Publicly Owned Electric Utilities without Generation Facilities by State (EIA-412)6	
www$financewogen_annual.xls <- "https://www.eia.gov/electricity/data/state/financewogen_annual.xls"

# Monthly data from Electric Power Monthly ####
# Electric Power Monthly with data for November 2019
#   Release date: January 30, 2020
#   Next release date: End of February 2020
  
# 2001 - Present	Net Generation by State by Type of Producer by Energy Source1	
www$generation_monthly.xlsx <- "https://www.eia.gov/electricity/data/state/generation_monthly.xlsx"
# 2001 - Present	Fossil Fuel Consumption for Electricity Generation by Year, Industry Type and State2	
www$consumption_monthly.xlsx <- "https://www.eia.gov/electricity/data/state/consumption_monthly.xlsx"}

sapply(www, download, overwrite = F)

# Annual capacity
elc_cap <- read_excel(file.path(eiadir, basename(www$existcapacity_annual.xlsx)),
                      sheet = "Existing Capacity",
                      range = "A2:H44621")
elc_cap
       
# Fuels consumption for electricity generation by months
fue4elc <- read_excel(file.path(eiadir, basename(www$consumption_monthly.xlsx)),
                      sheet = "2018_Preliminary",
                      range = "A5:F9361")
fue4elc

# Electricity generation by months and energy type
elc_gen <- read_excel(file.path(eiadir, basename(www$generation_monthly.xlsx)),
                      sheet = "2018_Preliminary",
                      range = "A5:F25156")
elc_gen

elc_emis <- read_excel(file.path(eiadir, basename(www$emission_annual.xls)),
                      sheet = "State Emissions",
                      range = "A1:G41765")
elc_emis

save(elc_cap, fue4elc, elc_gen, elc_emis, file = file.path(eiadir, "eia_raw.RData"))

```


## Power plants & generator level data  
### Form EIA-860 detailed data with previous form data (EIA-860A/860B)  
<https://www.eia.gov/electricity/data/eia860/> 
  Release date: September 3, 2019, Final 2018 data
  Next release date: June 2020, Early release 2019 data  
The survey Form EIA-860 collects generator-level specific information about existing and planned generators and associated environmental equipment at electric power plants with 1 megawatt or greater of combined nameplate capacity.  

### Form EIA-923 detailed data with previous form data (EIA-906/920)  
<https://www.eia.gov/electricity/data/eia923/>  
Monthly (M) release date: January 30, 2020 for November 2019 data
  Next monthly release: End of February 2020 (December 2019 data)
  Annual release date: September 20, 2019, Final 2018 data
  2018 Re-released: January 8, 2020  Corrections/Revisions  

The survey Form EIA-923 collects detailed electric power data -- monthly and annually -- on electricity generation, fuel consumption, fossil fuel stocks, and receipts at the power plant and prime mover level. Specific survey information provided:
* Schedule 2 - fuel receipts and costs  
* Schedules 3A & 5A - generator data including generation, fuel consumption and stocks  
* Schedule 4 - fossil fuel stocks  
* Schedules 6 & 7 - non-utility source and disposition of electricity  
* Schedules 8A-F - environmental data  
Monthly data (M) -approximately 2,026 plants from the monthly survey  
Annual final data - approximately 2,026 monthly plants + 7,278 plants from the annual survey   

Summary data  
<https://www.eia.gov/electricity/data.php#gencapacity>  

### Scraping & downloading all zip and xls files of the forms
(Another option is EIA's API <https://www.eia.gov/opendata/register.php>, though it is not clear yet if the API has all the the data in xls, need to check.)  
```{r EIA_860_923_forms}
scrap <- function(weblink, filepatern = "([a-zA-Z0-9_]+.zip)|([a-zA-Z0-9_]+.xls)$", aslist = T) {
  # library(rvest)
  links <- read_html(weblink) %>%
    html_nodes("a") %>%
    html_attr('href') 
  links <- links[grepl(filepatern, links)]
  links <- paste0(weblink, links)
  if (aslist) {
    lnames <- str_extract(links, filepatern)
    stopifnot(length(links) == length(lnames))
    links <- as.list(links)
    names(links) <- lnames
  }
  links
}

www <- scrap("https://www.eia.gov/electricity/data/eia860/")
www <- c(www, scrap("https://www.eia.gov/electricity/data/eia923/"))
sapply(www, download, overwrite = F)

fls <- names(www)

# Select data for 2018
fls <- fls[grepl("2018", fls)]
zz <- grepl("zip$", fls) # select zip-files for unzipping
# Unzip files
for (f in fls[zz]) {
  dirname <- gsub("\\.zip$", "", f)
  dirname <- file.path(eiadir, dirname)
  if (!dir.exists(dirname)) dir.create(dirname)
  unzip(file.path(eiadir, f), exdir = dirname, overwrite = T)
}

# dirname <- gsub("\\.zip$", "", fls[1])

```

Reading xls(x) files  
Note: reasing from Excel files should be supervised and adjusted in case of any updates of the files on EIA website. Also may not work as expected on some OS/platforms.
```{r}
elc_pp <- cleanread_xls(file.path(eiadir, "eia8602018/2___Plant_Y2018.xlsx"),
                      sheet = "Plant",
                      range = "A2:AP10982")
col_type <- sapply(elc_pp, class)
unique(col_type)
names(elc_pp)
unique(elc_pp$`Liquefied Natural Gas Storage`)

```

```{r, results="markdown", fig.cap="Mapping all power plants."}
load("data/maps/usa49reg.RData")
plot(usa49reg)
points(elc_pp$Longitude, elc_pp$Latitude, pch = 16, col = "blue", cex = .3)
save(elc_pp, file = file.path(eiadir, "elc_pp.RData"))

```


```{r}
# continue..
elc_gn <- cleanread_xls(file.path(eiadir, "eia8602018/3_1_Generator_Y2018.xlsx"),
                      sheet = "Operable",
                      range = "A2:BU22120")
col_type <- sapply(elc_gn, class)
unique(col_type)
elc_gn
names(elc_gn)

elc_genfue <- cleanread_xls(file.path(eiadir, "f923_2018/EIA923_Schedules_2_3_4_5_M_12_2018_Final_Revision.xlsx"),
                      sheet = "Page 1 Generation and Fuel Data",
                      range = "A6:CS13968")
elc_genfue
names(elc_genfue)
col_type <- sapply(elc_genfue, class)
unique(col_type)

# Split annual and monthly data
month_cols <- grepl(paste(month.name, collapse = "|"), names(elc_genfue))
summary(month_cols)

elc_genfue_y <- elc_genfue[, !month_cols]
dim(elc_genfue_y)

jj <- grepl("(Plant.id)|(YEAR)", names(elc_genfue), ignore.case = T)
elc_genfue_m <- bind_cols(elc_genfue[, which(jj)], elc_genfue[, month_cols])
dim(elc_genfue_m)

# reshape monthly data
aa <- gather(elc_genfue_m, key = "key", value = "value", -`Plant Id`, -YEAR)
aa$variable <- str_extract(aa$key, "^[a-zA-Z0-9_]+")
unique(aa$variable)
aa$month <- str_extract(aa$key, "[a-zA-Z0-9_]+$")
unique(aa$month)

elc_genfue_m <- select(aa, `Plant Id`, YEAR, month, variable, value)

elc_genfue <- left_join(elc_genfue_m, select(elc_genfue_y, -starts_with("Reserved")))

elc_genfue <- clean_names(elc_genfue)
names(elc_genfue)

```


### Processing plant-level data
```{r, eval=FALSE}
# Fuels
elc_genfue_y <- clean_names(elc_genfue_y)
names(elc_genfue_y)

unique(elc_genfue_y$Reported_Fuel_Type_Code)
unique(elc_genfue_y$AER_Fuel_Type_Code)

unique(elc_gn$Technology)
elc_gn <- clean_names(elc_gn)
names(elc_gn)

aa <- elc_gn %>% group_by(Technology, State) %>%
        summarise(cap = sum(`Nameplate_Capacity_MW`))

summary(aa)
sum(aa$cap)/1e3

(unique(aa$Technology))

```

# EPA data

## EPA's Power Sector Modeling Platform v6 - November 2018 Reference Case

"March 21, 2019 - EPA is making the latest power sector modeling platform available, including the associated input data and modeling assumptions, outputs, and documentation." <https://www.epa.gov/airmarkets/documentation-epas-power-sector-modeling-platform-v6-november-2018-reference-case>
Accessed: June 13, 2019

### Get IPM data files
```{r}
epadir <- file.path(getwd(), "data/EPA")
ipmdir <- file.path(epadir, "IPM")
# needsdir <- file.path(epadir, "IPM")

if(!dir.exists(ipmdir)) dir.create(ipmdir, recursive = T)

ipmweb <- "https://www.epa.gov/airmarkets/documentation-epas-power-sector-modeling-platform-v6-november-2018-reference-case"

scrap_ipm <- function(weblink, filepatern = "([a-zA-Z0-9_]+.zip)|([a-zA-Z0-9_]+.xls(|x))$", aslist = T) {
  # library(rvest)
  links <- read_html(weblink) %>%
    html_nodes("body") %>% html_nodes("section") %>% 
    html_nodes("div") %>% html_nodes("div") %>% 
    html_nodes("div") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes('span') %>% 
    html_nodes('a') %>% html_attr('href') 
  links <- links[grepl(filepatern, links)]
  links <- paste0("https://www.epa.gov", links)
  if (aslist) {
    lnames <- str_extract(links, filepatern)
    stopifnot(length(links) == length(lnames))
    links <- as.list(links)
    names(links) <- lnames
  }
  links
}
www <- scrap_ipm(ipmweb)
sapply(www, download, overwrite = F, path = ipmdir)

# IPM shape files
download(url = "https://www.epa.gov/sites/production/files/2019-08/ipm_v6_regions.zip",
         overwrite = F, path = ipmdir)
unzip(file.path(ipmdir, "ipm_v6_regions.zip"), 
      exdir = file.path(ipmdir, "ipm_v6_regions"), overwrite = T)

```


### Coal
```{r read_coal_data}
# Documentation
download(url = "https://www.epa.gov/sites/production/files/2019-03/documents/chapter_7.pdf", 
         path = ipmdir, overwrite = F)


# Read coal supply data
coa_sup <- cleanread_xls(file.path(ipmdir, "table_7-26_coal_supply_curves_in_epa_platform_v6.xlsx"),
                      sheet = "Table 7-26 Coal",
                      range = "B4:I9002")
col_type <- sapply(coa_sup, class)
unique(col_type)

coa_sup <- clean_names(coa_sup)
names(coa_sup)
coa_sup

unique(coa_sup$Coal_Grade)
unique(coa_sup$Coal_Supply_Region)
unique(coa_sup$Year)

# Coal type
coa_sup$Coal_Type <- substr(coa_sup$Coal_Grade, 1, 1)
unique(coa_sup$Coal_Type)

# Aggregate by coal types
coa_sup_agg <- coa_sup %>%
  group_by(Year, Coal_Type, Coal_Supply_Region) %>%
  summarise(
    Heat_Content_MMBtu_Ton = mean(Heat_Content_MMBtu_Ton, na.rm = T),
    Cost_of_Production_2016USD_Ton = 
      weighted.mean(Cost_of_Production_2016USD_Ton, 
                    Coal_Production_Million_Tons_Year, na.rm = T),
    Coal_Production_Million_Tons_Year = sum(Coal_Production_Million_Tons_Year, na.rm = T),
    Coal_Reserves_Million_Tons = sum(Coal_Reserves_Million_Tons, na.rm = T)
  )

coa_type_names <- c(B = "Bituminous", S = "Subbituminous", L = "Lignite")
coa_sup_agg$Coal_Type_Long <- plyr::revalue(coa_sup_agg$Coal_Type, coa_type_names)
save(coa_sup_agg, coa_sup, file = file.path(ipmdir, "coal_supply.RData"))
# write_csv(coa_sup_agg, path = "tmp/coa_sup_agg.csv")
# write_tsv(coa_sup_agg, path = "tmp/coa_sup_agg.tab")

```

### Gas
```{r read_gas_Data}
download(url = "https://www.epa.gov/sites/production/files/2019-03/documents/chapter_8.pdf", 
         path = ipmdir, overwrite = F)

# Read gas supply curves data
gas_sup_curves <- cleanread_xls(
  file.path(ipmdir, "table_8-5_natural_gas_supply_curves_for_epa_platform_v6.xlsx"),
  sheet = 1,
  range = "B4:D484")

gas_sup_curves <- clean_names(gas_sup_curves)

# Aggregated supply with average prices
gas_sup_avr <- gas_sup_curves %>%
  group_by(Year) %>%
  summarise(Price_2016USD = weighted.mean(Price_2016USD, Natural_Gas_Supply_to_Power_Sector_Quads),
            Natural_Gas_Supply_to_Power_Sector_Quads = sum(Natural_Gas_Supply_to_Power_Sector_Quads))

# Read basis data
gas_sup_basis <- cleanread_xls(
  file.path(ipmdir, "table_8-4_natural_gas_basis_for_epa_platform_v6.xlsx"),
  sheet = 1,
  range = "B4:F596")
gas_sup_basis <- clean_names(gas_sup_basis)

all(unique(gas_sup_basis$Year) == unique(gas_sup_avr$Year)) # Check

# Merge with prices
gas_sup_byreg <- full_join(gas_sup_basis, gas_sup_avr)
unique(gas_sup_byreg$IPM_Region)

save(gas_sup_avr, gas_sup_byreg, gas_sup_curves, gas_sup_basis,
     file = file.path(ipmdir, "gas_supply.RData"))

# write_tsv(gas_sup_avr, path = "tmp/gas_sup_avr.tab")
# write_tsv(gas_sup_byreg, path = "tmp/gas_sup_byreg.tab")


```


### NEEDS v6 Database
National Electric Energy Data System (NEEDS) v6
<https://www.epa.gov/airmarkets/national-electric-energy-data-system-needs-v6>
Accessed: Apr 28, 2020

The National Electric Energy Data System or “NEEDS” database contains the generation unit records used to construct the "model" plants that represent existing and planned/committed units in EPA modeling applications of IPM. NEEDS includes basic geographic, operating, air emissions, and other data on these generating units. NEEDS v6 is a complete bottom-up update for EPA's new power sector modeling platform v6.  For a description of the sources used in preparing NEEDS v6, see Documentation, Chapter 4: Generating Resources.


```{r}
needsweb <- "https://www.epa.gov/airmarkets/national-electric-energy-data-system-needs-v6"

# Most Recent Version of NEEDS:
# NEEDS v6 rev: 3-26-2020(5 MB)
# Comment and Change Log rev: 3-26-2020(142 K)
download(url = "https://www.epa.gov/sites/production/files/2020-04/needs_v6_03-26-2020.xlsx", 
         path = ipmdir, overwrite = F)


# NEEDS v6 Database(4 MB)  (rev: 5-31-2019)
# download(url = "https://www.epa.gov/sites/production/files/2019-06/needs_v6_may_2019_reference_case.xlsx", 
#          path = ipmdir, overwrite = F)


# Comment and Change Log(88 K) (rev: 5-31-2019)

# NEEDS v6 Database(3 MB)  (rev: 11-30-2018)
# download(url = "https://www.epa.gov/sites/production/files/2019-03/needs_v6_november_2018_reference_case_0.xlsx", 
#          path = ipmdir, overwrite = F)

# Incremental Documentation for NOx Rates in NEEDS v6(3 pp, 44 K) (9-14-2018)

# NEEDS v6 Database(3 MB) (rev: 5-30-2018)


scrap_needs <- function(weblink, filepatern = "([a-zA-Z0-9_]+.zip)|([a-zA-Z0-9_]+.xls(|x))$", aslist = T) {
  # library(rvest)
  links <- read_html(weblink) %>%
    html_nodes("body") %>% html_nodes("section") %>% 
    html_nodes("div") %>% html_nodes("div") %>% 
    html_nodes("div") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes("ul") %>% 
    html_nodes("li") %>% html_nodes('span') %>% 
    html_nodes('a') %>% html_attr('href') 
  links <- links[grepl(filepatern, links)]
  links <- paste0("https://www.epa.gov", links)
  if (aslist) {
    lnames <- str_extract(links, filepatern)
    stopifnot(length(links) == length(lnames))
    links <- as.list(links)
    names(links) <- lnames
  }
  links
}
www <- scrap_needs(needsweb)
sapply(www, download, overwrite = F, path = ipmdir)


```

```{r}
needs <- cleanread_xls(file.path(ipmdir, "needs_v6_03-26-2020.xlsx"),
                      sheet = "NEEDS v6_active",
                      range = "A1:AX18252")
col_type <- sapply(needs, class)
unique(col_type)
needs
names(needs)
save(needs, file = file.path(ipmdir, "NEEDS_v6_active.RData"))

```


### Wind potential
Table 4-38 Onshore Regional Potential Wind Capacity (MW) by TRG and Cost Class in EPA Platform v6
```{r}
tbl_4_38_pth <- file.path(
  ipmdir,
  "table_4-38_onshore_regional_potential_wind_capacity_mw_by_trg_and_cost_class_in_epa_platform_v6.xlsx")
file.exists(tbl_4_38_pth)

rename_fill_tbl <- function(tbl) {
  nms <- names(tbl)
  nms <- c(nms[1:3], paste0("Cost_Class_", 1:(ncol(tbl)-3)))
  names(tbl) <- nms
  tbl <- tbl[-1,]
  
  # fill-in NAs
  r_1 <- NA; s_1 <- NA
  for(i in 1:nrow(tbl)) {
    r <- tbl$IPM_Region[i]
    s <- tbl$State[i]
    if (!is.na(r)) r_1 <- r else tbl$IPM_Region[i] <- r_1
    if (!is.na(s)) s_1 <- s else tbl$State[i] <- s_1
  }
  tbl
}

reshape_MW_tbl <- function(tbl, val_to_name = "MW") {
  tbl <- tbl %>% 
  pivot_longer(cols = starts_with("Cost_Class"), names_prefix = "Cost_Class_", 
               names_to = "Cost_Class", values_to = val_to_name) %>%
  mutate(Cost_Class = as.integer(Cost_Class))
  tbl
}

onshore_wind_potential <- cleanread_xls(tbl_4_38_pth, "Table 4-38", skip = 3) %>%
  clean_names() %>% rename_fill_tbl() %>% reshape_MW_tbl()
onshore_wind_potential

save(onshore_wind_potential,
     file = file.path(ipmdir, "onshore_wind_potential.RData"))
rm(tbl_4_38_pth, onshore_wind_potential)

```


### Wind Generation Profiles 
in EPA Platform v6 (kWh of Generation per MW of Capacity)
```{r}
tbl_4_39_pth <- file.path(
  ipmdir, "table_4-39_wind_generation_profiles_in_epa_platform_v6",
  "Table 4-39 Wind Generation Profiles in EPA Platform v6.xlsx")
file.exists(tbl_4_39_pth)
# a <- read_excel(tbl_4_39_pth, "Onshore", skip = 2, col_names = T)

reshape_genpro_tbl <- function(x) {
  pivot_longer(x, cols = starts_with("Hour"), names_prefix = "Hour", 
               names_to = "hour", values_to = "kWh_per_MW") %>%
  mutate(hour = as.integer(hour) - 1L)
}

wind_onshore <- cleanread_xls(tbl_4_39_pth, "Onshore", skip = 3) %>%
  clean_names() %>% reshape_genpro_tbl()

wind_offshore_shallow <- cleanread_xls(
  tbl_4_39_pth, "Offshore Shallow", skip = 3) %>%
  clean_names() %>% reshape_genpro_tbl()

wind_offshore_middepth <- cleanread_xls(
  tbl_4_39_pth, "Offshore Mid-Depth", skip = 3) %>%
  clean_names() %>% reshape_genpro_tbl()

wind_offshore_deep <- cleanread_xls(
  tbl_4_39_pth, "Offshore Deep", skip = 3) %>%
  clean_names() %>% reshape_genpro_tbl()

save(wind_onshore, wind_offshore_shallow, 
     wind_offshore_middepth, wind_offshore_deep,
     file = file.path(ipmdir, "wind_profiles.RData"))
rm(tbl_4_39_pth, wind_onshore, wind_offshore_shallow, 
   wind_offshore_middepth, wind_offshore_deep)

```

### Wind Capital Cost Adder
Table 4-40 Capital Cost Adder (2016$/kW) for New Onshore Wind Plants by Resource and Cost Class in EPA Platform v6
```{r}
tbl_4_40_pth <- file.path(
  ipmdir,
  "table_4-40_capital_cost_adder_for_new_onshore_wind_plants_in_epa_platform_v6.xlsx")
file.exists(tbl_4_40_pth)

onshore_invcost_adder <- cleanread_xls(tbl_4_40_pth, "Table 4-40", skip = 3) %>%
  clean_names() %>% rename_fill_tbl() %>% 
  reshape_MW_tbl(val_to_name = "USD_per_kW")
onshore_invcost_adder  

save(onshore_invcost_adder,
     file = file.path(ipmdir, "onshore_invcost_adder.RData"))
rm(tbl_4_40_pth, onshore_invcost_adder)

```

### Solar potential
#### Photovoltaic
Table 4-41 Solar Photovoltaic Regional Potential Capacity (MW) by Resource and Cost Class in EPA Platform v6
```{r}
tbl_4_41_pth <- file.path(
  ipmdir,
  "table_4-41_solar_photovoltaic_regional_potential_capacity_mw_by_resource_and_cost_class_in_epa_platform_v6.xlsx")

if (file.exists(tbl_4_41_pth)) {
  solar_pv_potential <- cleanread_xls(tbl_4_41_pth, "Table 4-41", skip = 3) %>%
    clean_names() %>% rename_fill_tbl() %>% reshape_MW_tbl()
  
  save(solar_pv_potential,
       file = file.path(ipmdir, "solar_pv_potential.RData"))
  rm(tbl_4_41_pth, solar_pv_potential)
}

```

#### Solar Thermal
Table 4-42 Solar Thermal Regional Potential Capacity (MW) by Resource and Cost Class in EPA Platform v6
```{r}
tbl_4_42_pth <- file.path(
  ipmdir,
  "table_4-42_solar_thermal_regional_potential_capacity_mw_by_resource_and_cost_class_in_epa_platform_v6.xlsx")

if (file.exists(tbl_4_42_pth)) {
  solar_th_potential <- cleanread_xls(tbl_4_42_pth, "Table 4-42", skip = 3) %>%
  clean_names() %>% rename_fill_tbl() %>% reshape_MW_tbl()
  
  save(solar_th_potential,
       file = file.path(ipmdir, "solar_th_potential.RData"))
  rm(tbl_4_42_pth, solar_th_potential)
}

```

### Solar generation profiles
Table 4-43 Solar Photovoltaic Generation Profiles in EPA Platform v6 (kWh of Generation per MW of Capacity)
```{r}
tbl_4_43_pth <- file.path(
  ipmdir,
  "table_4-43_solar_photovoltaic_generation_profiles_in_epa_platform_v6.xlsx")

if (file.exists(tbl_4_43_pth)) {
  solar_profiles <- cleanread_xls(tbl_4_43_pth, "Table 4-43", skip = 3) %>%
    clean_names() %>% reshape_genpro_tbl()
  solar_profiles
  
  save(solar_profiles, file = file.path(ipmdir, "solar_profiles.RData"))
  rm(tbl_4_43_pth, solar_profiles)
}

```

### Solar Capital costs adder
#### Photovoltaic
Table 4-44 Solar Photovoltaic Regional Capital Cost Adder (2016$/kW) for Potential  Units by Resource and Cost Class in EPA Platform v6
```{r}
tbl_4_44_pth <- file.path(
  ipmdir,
  "table_4-44_capital_cost_adder_for_new_solar_pv_plants_in_epa_platform_v6.xlsx")
file.exists(tbl_4_44_pth)

solpv_invcost_adder <- cleanread_xls(tbl_4_44_pth, "Table 4-44", skip = 3) %>%
  clean_names() %>% rename_fill_tbl() %>% 
  reshape_MW_tbl(val_to_name = "USD_per_kW")
solpv_invcost_adder  

save(solpv_invcost_adder,
     file = file.path(ipmdir, "solpv_invcost_adder.RData"))
rm(tbl_4_44_pth, solpv_invcost_adder)

```

#### Solar Thermal
Table 4-45 Solar Thermal Regional Capital Cost Adder (2016$/kW) for Potential Units by Resource and Cost Class in EPA Platform v6
```{r}
tbl_4_45_pth <- file.path(
  ipmdir,
  "table_4-45_capital_cost_adder_for_new_solar_thermal_plants_in_epa_platform_v6.xlsx")
file.exists(tbl_4_45_pth)

solth_invcost_adder <- cleanread_xls(tbl_4_45_pth, "Table 4-45", skip = 3) %>%
  clean_names() %>% rename_fill_tbl() %>% 
  reshape_MW_tbl(val_to_name = "USD_per_kW")
solth_invcost_adder  

save(solth_invcost_adder,
     file = file.path(ipmdir, "solth_invcost_adder.RData"))
rm(tbl_4_45_pth, solth_invcost_adder)

```


### Load duration curves
Table 2-2 2021 Load Duration Curves Used in EPA Platform v6
```{r}
tbl_2_2_pth <- file.path(
  ipmdir,
  "table_2-2_load_duration_curves_used_in_epa_platform_v6.xlsx")
file.exists(tbl_2_2_pth)

load_curves <- cleanread_xls(tbl_2_2_pth, "Table 2-2", skip = 3) %>%
  clean_names() %>%
  pivot_longer(cols = starts_with("Hour"), names_prefix = "Hour_", 
               names_to = "hour", values_to = "MWh") %>%
  mutate(hour = as.integer(hour) - 1L)

save(load_curves, file = file.path(ipmdir, "load_curves.RData"))
rm(tbl_2_2_pth, load_curves)

```

### Transmission capabilities  
Table 3-21 Annual Transmission Capabilities of U.S. Model Regions in EPA Platform v6 - 2021
```{r}
tbl_3_21_pth <- file.path(
  ipmdir,
  "table_3-21_annual_transmission_capabilities_of_u.s._model_regions_in_epa_platform_v6_-_2021.xlsx")
file.exists(tbl_3_21_pth)

transmission <- cleanread_xls(tbl_3_21_pth, "Table 3-21", skip = 3) %>%
  clean_names()

# fill-in NAs in "from" column
transmission$From2 <- transmission$From
v_1 <- NA
for(i in 1:nrow(transmission)) {
  v <- transmission$From[i]
  if (!is.na(v)) v_1 <- v
  transmission$From2[i] <- v_1
}
save(transmission, file = file.path(ipmdir, "transmission.RData"))
rm(tbl_3_21_pth, transmission)

```


### eGRID
...

# NREL  
<https://atb.nrel.gov/electricity/2018/summary.html>  
<https://atb.nrel.gov/electricity/2019/data.html>
(included in PowerGenome database)  

# FERC  
<https://www.ferc.gov/docs-filing/forms/form-1/data.asp>    
(included in PUDL database)  

# PUDL  
[The Public Utility Data Liberation (PUDL) project](https://catalyst.coop/pudl/) is an assembly of EIA, EPA, and FERC forms, standardized and documented.  

# PowerGenome  
<https://github.com/gschivley/PowerGenome>  
A tool to create input files for power system optimization models with an integrated database (a selection from PUDL (EIA, EPA, FERC), NREL, and other sources).  
The script in the chunk below reads PowerGenome SQLite-database, re-saves it in R-format, also takes a glimpse at some tables in the database. To run the script, download the *pudl_updated.sqlite* (see the link at <https://github.com/gschivley/PowerGenome>). 

```{r, eval=FALSE, include=FALSE}
# Try PowerGenome 
if (file.exists("tmp/PowerGenome.RData")) {
  load("tmp/PowerGenome.RData")
} else {
  if (!dir.exists("data/PowerGenome/")) dir.create("data/PowerGenome/")
  if (!dir.exists("tmp/")) dir.create("tmp/")

  pg <- file.path("<PATH TO THE POWERGENOME DATABASE>/pudl_updated.sqlite")
  file.exists(pg)
  library(DBI)
  library("RSQLite")
  # Create an ephemeral in-memory RSQLite database
  con <- dbConnect(RSQLite::SQLite(), pg)
  con
  dbListTables(con)
  
  db2list <- function(db, tables = NULL, asTibble = TRUE, verbose = TRUE) {
    if (is.null(tables)) tables = dbListTables(db)
    t0 <- Sys.time()
    ll <- list()
    for (x in tables) {
      cat("reading ", x)
      y <- dbReadTable(db, x)
      cat(",", size(y), "\b,", round(Sys.time() - t0, 1), "\bs\n")
      if (asTibble) y <- as_tibble(y)
      ll[[x]] <- y
    }
    ll
  }

  db <- db2list(con, dbListTables(con))
  size(db)
  names(db)
  length(db)
  dbDisconnect(con)
  save(db, file = "tmp/PowerGenome.RData") # the data will not be redistributed
  
}

# IPM tables ####
names(db)[grepl("ipm", names(db), T)]

# Regions ####
unique(db$plant_region_map_epaipm$region)
unique(db$transmission_single_epaipm$region_from) # 78 IPM regions (67 US + 11 CA)
unique(db$load_curves_epaipm$region_id_epaipm)
unique(db$plants_entity_eia$nerc_region) # plant_id_eia, state, county, NERC, balancing, GIS
unique(db$regions_entity_epaipm$region_id_epaipm) # IMP 78 regions
reg_names_ipm78 <- db$regions_entity_epaipm$region_id_epaipm
save(reg_names_ipm78, file = "data/PowerGenome/regions_entity_epaipm.RData")

# Demand ####
## Load curves ###
db$load_curves_epaipm
unique(db$load_curves_epaipm$region_id_epaipm)
unique(db$load_curves_epaipm$month)
unique(db$load_curves_epaipm$hour)
unique(db$load_curves_epaipm$day_of_year)
length(unique(db$load_curves_epaipm$time_index))

# Commodities ####
db$fuel_type_aer_eia923 #18
db$energy_source_eia923 #20
db$fuel_type_eia923 #42
unique(db$generation_fuel_eia923$fuel_type)
unique(db$generation_fuel_eia923$fuel_type_code_pudl)
unique(db$generation_fuel_eia923$fuel_type_code_aer)
unique(db$fuel_receipts_costs_eia923$fuel_group_code)
unique(db$fuel_receipts_costs_eia923$fuel_group_code_simple)
unique(db$fuel_receipts_costs_eia923$energy_source_code)

# Technologies ####
## boilers and generators
db$prime_movers_eia923 # movers codes and names
db$boiler_fuel_eia923  %>% as.data.frame() %>% head() # fuels consumed by months
db$generation_eia923 # generation by months
db$generation_fuel_eia923 %>% as.data.frame() %>% head() # both by months
db$generators_eia860 %>% as.data.frame() %>% head() # capacity

## plants meta info
db$generators_entity_eia %>% as.data.frame() %>% head() # operating_date
db$plant_region_map_epaipm # plants_id_eia by IMP regions
db$plants_eia #  eia vs pudl
db$plants_eia860 %>% as.data.frame() %>% head() # plants techs detailes
db$plants_entity_eia %>% as.data.frame() %>% head() # GIS
db$plants_ferc1 # plants vs. utility
db$plants_hydro_ferc1 %>% as.data.frame() %>% head() # invcost
db$plants_pumped_storage_ferc1 %>% as.data.frame() %>% head() # invcost
db$plants_small_ferc1 %>% as.data.frame() %>% head() # invcost
db$plants_steam_ferc1 %>% as.data.frame() %>% head() # invcost
db$plants_pudl
## New techs ####
db$technology_costs_nrelatb %>% as.data.frame() %>% head()
db$technology_heat_rates_nrelatb

## Emissions ####
db$hourly_emissions_epacems %>% as.data.frame() %>% head()
summary(db$hourly_emissions_epacems$unit_id_epa)
summary(is.na(db$hourly_emissions_epacems$facility_id))

## IDs ####
db$boiler_generator_assn_eia860
db$boilers_entity_eia
unique(db$generators_eia860$technology_description)

# Supply ####
head(db$fuel_ferc1) %>% as.data.frame()
head(db$fuel_receipts_costs_eia923) %>% as.data.frame()
db$transport_modes_eia923

# Transmission ####
db$transmission_single_epaipm
unique(db$transmission_single_epaipm$region_from) # 78 regions 
unique(db$transmission_single_epaipm$region_to)

# Misc ####
db$ferc_depreciation_lines %>% as.data.frame()
db$coalmine_eia923
db$ferc_accounts
db$ownership_eia860 %>% as.data.frame() %>% head()
db$plant_in_service_ferc1 %>% as.data.frame() %>% head()
db$purchased_power_ferc1 %>% as.data.frame() %>% head()
## Utilities
db$utilities_eia
db$utilities_eia860
db$utilities_entity_eia
db$utilities_ferc1
db$utilities_pudl
db$utility_plant_assn

```

