---
title: "US 26-zone minimal model for testing"
subtitle: "A reproduction of MIP-project model with aggregated resources"
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(usensys)
```

```{r, include=FALSE}
# Load PowerGenome data
# 52-week hourly data for 26 zones
# CO2-cap, "base" scenario
data_dir <- "Z:/MIP/20240515/base_52_week"

```

## Model characteristics

The model is based on the 26-zone model from the [Multi-model
intercomparison project](link). The dataset is prepared by the
PowerGenome and Zero Lab teams. It includes 52 weeks of hourly data for
26 zones. Since the goal of the MIP project is to compare different
models, the model setting has a most basic, most common for capacity
expansion models features to simplify the comparative process across
models. Here are the main characteristics of the MIP model:

-   26 zones\
-   52 weeks with hourly resolution\
-   2027 - 2050 time horizon, 2027 as the base year, milestones every 5
    years starting from 2030\
-   clustered generation technologies and variable capacity factors for
    wind and solar generators\
-   NREL's Energy Technology Perspectives (ETP) projections of
    technologies and costs (included in PowerGenome), median scenario is
    used for the project\
-   age-based retirement assumptions of existing capacity, early
    retirement is allowed in some scenarios\
-   current policy constraints, renewable portfolio standards,
    deployment plans for renewable energy sources, regional CO2 cap\
-   national CO2 cap with rapid decline after 2030.

Combinations of these features are used to create different scenarios
for the project. The reproduction of the full model with all scenarios
will be published in a separate repository. This document focuses on the
minimal model with aggregated resources for demonstration and testing
purposes.

The key difference between the model used in MIP project and the
"minimal" model currently available in the repository is in aggregation
of resources with a goal to reduce the computational time and the size
of the dataset. Variable capacity factors for wind and solar generators
are clustered within zones based on correlation of time-series data. The
clustering algorithm insures that the aggregated capacity factors are as
close as possible to the original data with maximum loss of variability
up to 5%. Thermal generation is also gropped by fuel type with averaging
of technical and economic parameters.
